---
title: "Assignment 2 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "16/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 2: meta-analysis

## Questions to be answered

1. Simulate data to setup the analysis and gain insight on the structure of the problem. Simulate one dataset of 100 studies (n of participants should follow a normal distribution with mean of 20, sd of 10, but no fewer than 10 participants), with a mean effect size of 0.4, average deviation by study of .4 and measurement error of .8. The data you get should have one row per study, with an effect size mean and standard error. Build a proper bayesian model to analyze the simulated data. Then simulate publication bias (only some of the studies you simulate are likely to be published, which?), the effect of publication bias on your estimates (re-run the model on published studies, assess the difference), and discuss what this implies for your model. remember to use at least one plot to visualize your results. 
BONUS question: do a power/precision analysis: w this kind of sample sizes (participants) how many studies would you need to acquire good precision (e.g. .1 sd in the pop level estimate)

2. What is the current evidence for distinctive vocal patterns in schizophrenia? 
Use the data from Parola et al (2020) - https://www.dropbox.com/s/0l9ur0gaabr80a8/Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx?dl=0 - focusing on pitch variability (PITCH_F0SD).  Describe the data available (studies, participants). Using the model from question 1 analyze the data, visualize and report the findings: population level effect size; how well studies reflect it; influential studies, publication bias. 
BONUS question: assess the effect of task on the estimates (model comparison with baseline model)


# Question 1

```{r}
remove.packages(c("Rcpp","dplyr","rlang"))
install.packages(c("Rcpp","dplyr","rlang"))
library(tidyverse)
library(brms)
library(gridExtra)
library(cmdstanr)
library(tidymodels)
library(truncnorm)
```
### 1.a Simulate DATA

```{r}
set.seed(3)
#making a df
sim_meta_df <- tibble( ID = rep(1:100),
                       participants = as.integer(rtruncnorm(100, mean = 20, sd= 10, a = 10)),
                       effect_no_error = rnorm(100,0.4,0.4),
                       effectmu = NA,
                       effectsigma = NA
  
)
# i don´t know how to do it in tidyverse so base R it is
for (i in 1:100){

sampling <- rnorm(sim_meta_df$participants[i],sim_meta_df$effect_no_error[i],.8)

sim_meta_df$effectmu[i] <- mean(sampling)
sim_meta_df$effectsigma[i] <- sd(sampling)/sqrt(sim_meta_df$participants[i])
}
```
### 1.b Bayesian model


## Question 2

```{r}
set.seed(3)
#model
sim_formula <- effectmu | se(effectsigma) ~ 1 + (1|ID)
#priors
#get_prior(sim_formula,sim_meta_df)

sim_prior <- c(
  prior(normal(0,.3), class = Intercept),
  prior(normal(0,.3), class = sd)
)
#pp check

sim_m1_pp <- brm(
  sim_formula,
  data= sim_meta_df,
  prior = sim_prior,
  sample_prior = "only",
  backend = "cmdstanr",
  chains = 2,
  cores = 5,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
  ))
pp_check(sim_m1_pp,ndraws = 100)

#fit
sim_m1 <- brm(
  sim_formula,
  data= sim_meta_df,
  prior = sim_prior,
  sample_prior = T,
  backend = "cmdstanr",
  chains = 2,
  cores = 5,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
  ))

#pp check 2
pp_check(sim_m1,ndraws = 100)
#update plots
post_sim_m1 <- as_draws_df(sim_m1)

grid.arrange(
ggplot(post_sim_m1) +
  geom_density(aes(prior_Intercept), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('Intercept') +
  theme_classic(),

ggplot(post_sim_m1) +
  geom_density(aes(prior_sd_ID), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(sd_ID__Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('sd') +
  theme_classic()
)

#summary
print(sim_m1)

#visualize
sim_m1 %>% 
  fixef() %>% #takes estimates from model
  as.data.frame() %>% #datarames the matrix
  rownames_to_column("Betas") %>% #makes the first column a data frame column
  ggplot(aes(x=Betas))+
         geom_pointrange(fatten = 1/2,aes(y = Estimate, ymin =Q2.5,ymax = Q97.5))+
  geom_abline(linetype = 2,intercept = 0,slope = 0)+
  ylim(-0.15,0.5)+
  theme_classic()

```

### 1.c adding publication bias

```{r}
#subsetting based on whether the effect mean is at least twice the size of standard error (scientific practice)
set.seed(3)
sim_meta_bias_df <- sim_meta_df 

for (i in 1:100) {
  
  sim_meta_bias_df$published[i] <- ifelse(abs(sim_meta_bias_df$effectmu[i]) - 
                                            2*sim_meta_bias_df$effectsigma[i]
                                          > 0, rbinom(1,1,.9), rbinom(1,1,.1))
}

sim_meta_bias_df <- sim_meta_bias_df %>% 
  filter(published == 1)

```

```{r}
#modeling with published studies

set.seed(3)
#pp check

sim_m1_bias_pp <- brm(
  sim_formula,
  data= sim_meta_bias_df,
  prior = sim_prior,
  sample_prior = "only",
  backend = "cmdstanr",
  chains = 2,
  cores = 5,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
  ))
pp_check(sim_m1_bias_pp,ndraws = 100)

#fit
sim_m1_bias <- brm(
  sim_formula,
  data= sim_meta_bias_df,
  prior = sim_prior,
  sample_prior = T,
  backend = "cmdstanr",
  chains = 2,
  cores = 5,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
  ))

#pp check 2
pp_check(sim_m1_bias,ndraws = 100)
#update plots
post_sim_m1_bias <- as_draws_df(sim_m1_bias)

grid.arrange(
ggplot(post_sim_m1_bias) +
  geom_density(aes(prior_Intercept), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('Intercept') +
  theme_classic(),

ggplot(post_sim_m1_bias) +
  geom_density(aes(prior_sd_ID), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(sd_ID__Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  xlab('sd') +
  theme_classic()
)

#summary
print(sim_m1_bias)

#visualize differences 
diff_df <- rbind(sim_m1_bias %>% 
  fixef() %>% #takes estimates from model
  as.data.frame() %>% #datarames the matrix
  rownames_to_column("Betas"),
  sim_m1 %>% 
  fixef() %>% #takes estimates from model
  as.data.frame() %>% #datarames the matrix
  rownames_to_column("Betas"))
  #makes the first column a data frame column
diff_df[1,1] <- "Biased_Intercept"
diff_df %>% 
  ggplot(aes(x=Betas))+
         geom_pointrange(fatten = 1/2,aes(y = Estimate, ymin =Q2.5,ymax = Q97.5))+
  geom_abline(linetype = 2,intercept = 0,slope = 0)+
  ylim(-0.15,0.5)+
  theme_classic()


```

