---
title: "Untitled"
author: "Balazs Szabo"
date: "2022-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
remove.packages(c("Rcpp","dplyr","rlang"))
install.packages(c("Rcpp","dplyr","rlang"))
library(tidyverse)
library(brms)
library(gridExtra)
library(cmdstanr)
library(tidymodels)
library(truncnorm)
```
## Part I - Simulating data

Use the meta-analysis reported in Parola et al (2020), create a simulated dataset with 100 matched pairs of schizophrenia and controls, each participant producing 10 repeated measures (10 trials with their speech recorded). for each of these "recordings" (data points) produce 10 acoustic measures: 6 from the meta-analysis, 4 with just random noise. Do the same for a baseline dataset including only 10 noise variables. Tip: see the slides for the code. 

```{r}
#number of participants
n = 100

#effect means
Informed_effect_mean <- c(.25,-.55,-.75,-1.26,.05,1.89,0,0,0,0)
Skeptic_effect_mean <- rep(0,10)

#individual sds
Individual_sd <- 1
Trial_sd <- .5
Error <- .2
  
  
#simulate it 
inf_simdata <-
  tibble(
    ID = rep(rep(seq(1,n), each = 10),2),
    Diagnosis = rep(c("Schizophrenia","Control"), each = n*10),
    Trial = rep(seq(1,10), n*2)
  )


#figuring out a solution
 teszt <- tibble(
    ID = rep(rep(seq(1,n), each = 10),2),
    Diagnosis = rep(c("Schizophrenia","Control"), each = n*10),
    Trial = rep(seq(1,10), n*2)
  )
  #  the mapply function applies the rnorm to get one value, 100 times, eith the means of 3 different groups a 100 times. so it will return 100 random values for the 3 different means, but since we need 10 to be the same, because it is the "true" value for the participants, we rep each 10 times.
  teszt2 <- rep(mapply(rnorm, n = rep(1,100), mean = rep(c(0,100,1000), each = 100), sd = Individual_sd), each = 10)
  
  #the 3000 element vector then gets turned into a matrix with 3 1000 long columns
  teszt3 <- as.data.frame(matrix(teszt2,ncol =3,byrow = F))
  #because the true effect is the same for the two parts of the control group, we bind it with itself
  teszt4 <- rbind(teszt3,teszt3)
# we add the columns to the existing dataframe
teszt <- teszt %>% 
     tibble::add_column(!!!set_names(teszt4,
                                     nm=c("vt1","vt2","vt3")))



```
```{r}
# Ricardos code for reference

for (i in seq(10)){
  
  ric_temp <- tibble(
    ID = seq(n),
    Trueffect = rnorm(n, Informed_effect_mean[i], Individual_sd),
    Variale = paste0("v",i)
  )
  if (i==1){
    d_ric <- ric_temp
    }else{
      d_ric <- rbind(d_ric,ric_temp)
    }
  
}
```

```{r}
set.seed(3)
#continue the simulation

#  the mapply function applies the rnorm to get one value, 100 times, eith the means of 3 different groups a 100 times. so it will return 100 random values for the 10 different means, but since we need 10 to be the same, because it is the "true" value for the participants, we rep each 10 times.
  inf_help1 <- rep(mapply(rnorm, n = rep(1,100), mean = rep(Informed_effect_mean, each = 100), sd = Individual_sd), each = 10)
  
  #the 3000 element vector then gets turned into a matrix with 10,  1000 long columns
  inf_help2 <- as.data.frame(matrix(inf_help1,ncol =10,byrow = F))
  #because the true effect is the same for the two parts of the control group, we bind it with itself
  inf_help3 <- rbind(inf_help2,inf_help2)
# we add the columns to the existing dataframe
inf_simdata <- inf_simdata %>% 
     tibble::add_column(!!!set_names(inf_help3,
                                     nm=c("vt1","vt2","vt3","vt4","vt5",
                                          "vt6","vt7","vt8","vt9","vt10")))


```

